Tight Bounds for Dynamic Balanced Graph Partitioning

### 2020-05-10 paper/disc20clustering.tex
Typos
Big grammar check of all paper; several issues fixed

\begin{abstract}
	Distributed   applications,  including  batch  processing, streaming, scale-out databases,
	or machine learning, generate a significant amount of network traffic.
	By collocating frequently communicating nodes (e.g., virtual machines) on the same clusters (e.g., server or rack), we can reduce the network load and  improve application performance. 
	However, the communication pattern of different application is often unknown a priori, and may change over time, hence it needs to be learned in an online manner.
	%
	This paper revisits the online 
	balanced repartitioning problem 
	(introduced by Avin et al.~at DISC 2016)
	which asks for an algorithm that strikes
	an optimal tradeoff between the benefits
	of collocation (i.e., lower network load) 
	and its costs (i.e., migrations). 
	%
	Our first contribution is a significantly improved
	lower bound of $\Omega(k\cdot \ell)$ on the
	competitive ratio, where $\ell$ is the number
	of clusters and $k$ is the cluster size,
	even for a scenario in which the communication
	pattern is static and can be perfectly partitioned;
	we also provide a \mahmoud{asymptotically?} tight upper bound 
	of $O(k\cdot \ell)$ for this scenario.
	We contribute a tight upper bound
	of $\Theta(\ell)$ for $k=3$,
	for the general model in which the
	communication pattern can change arbitrarily
	over time.
	We improve the result for $k=2$ by providing a strictly $6$-competitive upper bound for the general model.
	In contrast to most prior work, our algorithms respect all capacity constraints, and do not require resource augmentation.

@@

The popularity of data-centric, distributed applications has led to an explosive growth of network traffic, especially in data centers \cite{roy2015inside,singh2015jupiter}.
As the performance of these distributed applications often critically depends on the underlying network \cite{mogul2012we}, an efficient operation of these networks is important.
At the same time, distributed systems are often highly virtualized today, and provide interesting new opportunities for resource optimization.
In particular, it has become possible to operate data centers in a more demand-aware manner: 
by dynamically migrating nodes (e.g., virtual machines) which communicate frequently topologically closer to each other, network traffic can be reduced significantly.  
However, migrations should be used moderately, as migrations also entail overheads.

@@

An algorithm serves a communication request between two nodes
either \emph{locally} at cost~0
if they are collocated,
or \emph{remotely} at cost~1
if they are located in different clusters.
We refer to these two types of requests as \emph{internal}
and \emph{external} requests, respectively.
Before serving a request,
an online algorithm may perform a \emph{repartition},
%(i.e., \emph{reconfigure}).
i.e.,
it may move (``migrate'') some nodes into clusters different from their current clusters, while respecting the capacity of every cluster. 
Afterwards, 
the algorithm serves the  request.
The cost of migrating a node from one cluster to another
is~$\alpha \in \mathbb{Z}^+$.
For any algorithm $\ALG$,
its cost,
denoted by $\ALG(\sigma)$,
is the total cost of communications and
the cost of migrations performed by $\ALG$ while serving the sequence $\sigma$.

@@

The two works closest to ours are by Avin et al. (on the general partitioning model) \cite{repartition-disc, sidma-arxiv} and by Henzinger et al. (on the learning model) \cite{sigmetrics19_partitioning}.
The focus of these papers however is primarily on models with resource augmentation: the online algorithm can use slightly larger clusters than the offline algorithm.  
Avin et al. actually showed that their lower bound $\Omega(k)$ holds even for a significant resource augmentation, and they provided an algorithm with the competitive ratio $O(k \log k)$ using the $(2+\epsilon)$-augmented cluster capacity.
Their ratio is independent of $\ell$, which is impossible without significant resource augmentation.

@@

In this section we consider the learning variant of Dynamic Balanced Graph Partitioning.
For this setting, we show a surprisingly high lower bound of $\Omega(k \cdot \ell)$.
The lower bound can be easily translated to the partitioning model (studied in Section~\ref{sec:part}) with finite request sequence.
At the end of this section, we discuss an asymptotically optimal upper bound.

@@

A~\emph{ground set} is a group of nodes that repeatedly communicate when an algorithm splits them (places at different clusters).
Their role is to ensure that the algorithm maintains a~partition of ground sets into clusters.
We emphasize that ground sets are revealed only upon their nodes are split.
We say that a ground set is \emph{trivial}, if it contains exactly one node.
We say that the node that belongs to trivial a ground set is \emph{isolated}.

The adversary constructs ground sets depending on choices of a deterministic online algorithm.
We emphasize that the algorithm does not know ground sets before they are revealed.
Once we construct a ground set, it lasts until the end of the input sequence.

@@

\begin{proof}
	Fix any online algorithm \ALG{}.
	Let $I(C)$ denote the cluster where nodes of the ground set $C$ are located in the initial configuration.
	If $C$ contains nodes that originate from different clusters, $I(C)$ is undefined.
	Throughout the construction, we maintain ground sets for each node.
	Initially all nodes are isolated (i.e., in ground sets of size $1$).
	First, we create a ground set $B$ of $k-1$ nodes on an arbitrarily chosen cluster (the nodes  are already collocated).
	%We never dismiss any constructed ground set.
	Each cluster hosts exactly $k$ nodes, and in any configuration a single isolated node resides in $S$.
	We say that this isolated node that accompanies the ground set $B$ is \emph{special}.
	We denote the first special node by $x_0$.

@@

	We claim that expanding ground sets in this way can continue as long as at least $\ell$ isolated nodes exists.
	Fix the configuration just before the creation of a ground set with $x_i$ for some $i$.
	We show that if at least $\ell$ isolated nodes exist (including $x_i$), then there exists a~perfect partition of ground sets after expanding a ground set with $x_i$.
	To show this claim, we consider a configuration, where the largest ground set of each cluster originates from it (excluding the ground set $\{x_0, y_0\}$).
	In this configuration, each cluster contains:  at most one ground set larger than $1$, possibly some isolated nodes, and possibly the ground set $\{x_0, y_0\}$.
	If $I(x_i)$ contains an isolated node, then we simply obtain a~perfect partition by swapping it with $x_i$.
	In the following we assume that no isolated nodes exist on $I(x_i)$.
	The largest ground set of $I(x_i)$ is at most $k-1$, because $x_i$ is on $S$.
	Therefore $\{x_0, y_0\}$ is present on $I(x_i)$, and we must evict it to make space for $x_i$.
	To find the cluster for $\{x_0, y_0\}$, observe that we have at most $\ell-1$ clusters with isolated nodes, and we have at least $\ell$ isolated nodes --- hence a~cluster with two isolated nodes exists.
	To find a feasible partition, we swap $\{x_0, y_0\}$ with these two collocated isolated nodes, and then we swap an isolated node from $I(x_i)$ with $x_i$.

@@

\noindent
\textbf{General partitioning problem.}
Every lower bound for the learning problem constitutes a converging lower bound for the partitioning problem as the input sequence grows to infinity.
However, our construction allows for an exact lower bound for partitioning with finite input sequence.
To this end, we issue only enough requests until the algorithm collocates the nodes of each ground set (if it never does, we terminate after it pays $k\cdot \ell$).
The construction is oblivious to the choice of the reconfiguration cost $\alpha$: we compare the 	number of node exchanges of \ALG and \OPT.

@@

Finally, we note the possibility for improvement. The algorithm CREP~\cite{repartition-disc} requires $(2+\epsilon)$-augmentation to guarantee the competitive ratio independent of $\ell$.

@@

We say that a component is \emph{trivial}, if it contains exactly one node,
and the node in this component is \emph{isolated}.
We say that a~partition is \emph{component-respecting}
if the nodes belonging to the same component are always collocated.

The algorithm in this section, and the algorithm for $k=3$ (cf. Section~\ref{sec:k3}) are modified versions of the algorithm DET from \cite{repartition-disc}.
The difference is in the the choice of partition after a component merge.
In DET, the partition was arbitrary.
In the algorithm for the learning model, we choose the partition closest to the initial partition.
\maciek{component respecting?}
In the algorithm from Section~\ref{sec:k3}, we choose the partition closest to the current partition (a repartition of minimum cost).

@@

With each repartitioning,
\PPL moves to a \maciek{component-respecting} partition that minimizes the distance to the initial partition $P_I$.
As a result,
\PPL never moves to a~partition that is more than $\Delta^*$  node migrations away from $P_I$.
This invariant latter ensures us that \PPL does not pay too much while recovering $P_F$.
We emphasis that a repartitioning by \PPL replaces the current partition $P$ with a~perfect partition closest to $P_I$.
Meaning, it never moves to a partition beyond distance $\Delta^*$.

@@

\begin{lemma}	\label{lemma:rebalancecost}
	The cost of each repartitioning by \PPL is $2\cdot\OPT$.
\end{lemma}
\begin{proof}
	Let $P_i$ denote the partition of \PPL immediately after serving $\sigma_{i}$.
	Consider the repartitioning that transforms 
	$P_{t-1}$ to $P_t$ upon the request $\sigma_t$.
	Let $M \subset V$ denote the set of nodes that migrate during this process.
	Let $M^-$ and $M^+$ denote the subset of nodes that, respectively,
	enter or leave their initial cluster during the repartitioning.    
	Then,
	$M = M^+ \cup M^-$.
	Since at least $|M^-|$ nodes are not in their initial cluster before the repartitioning (i.e., in $P_{t-1}$),
	the distance before the repartitioning is $\Delta(P_{t-1}) \geq | M^-|$.
	Analogously,
	the distance afterwards is $\Delta(P_{t}) \geq | M^+|$.
	Thus,
	$|M| \leq \Delta(P_{t-1}) + \Delta(P_{t})$.
	By Property \ref{prop:dist<OPT},
	$\Delta(P_{t-1})  \leq \Delta^*$ and
	$\Delta(P_{t}) \leq \Delta^*$.
	Since $\Delta^* \leq \OPT$,
	we obtain	 $|M| \leq 2\cdot\OPT$.
\end{proof}

@@

Now  we discuss the general online
model where the request sequence
can be arbitrary.
In Section~\ref{sec:k3}, we show a $O(k \cdot \ell)$-competitive algorithm for $k=3$ using the classic \emph{rent-or-buy} approach~\cite{karlin-ski-rental}.
Prior to this section, we showed a lower bound of $\Omega(k \cdot \ell)$  that holds for the general model, hence the result from this section is asymptotically optimal (cf. Section~\ref{sec:lowerbound}).
Furthermore, in Section~\ref{sec:k2}, we show a strictly $6$-competitive algorithm for $k=2$.

@@

	If either $U$ or $V$ is of type $C_1$, then this cluster can fit the merged component, and the reconfiguration is local within $U$ and $V$.
	In this case, a local repartition is possible (within $U$ and $V$), for a cost of at most $2$ swaps.
	Note that it is impossible that either $U$ or $V$ is of type $C_3$, as otherwise, a components of size $3$ participates in a merge.
	After merge, we would have a component of size at least $4$, and in this case no component respecting partitioning exists, a contradiction.

	It remains to consider the case where both $U$ and $V$ are of type $C_2$. Note that $(u,v)$ cannot both belong to a component of size $2$, as this would mean that \TAlg has the component of size $4$, a contradiction with the case assumption that the component respecting partitioning exists. 
	Otherwise, if either $u$ or $v$ belongs to a component of size $2$, then it suffices to exchange components of size $1$ between $U$ and $V$.
	Finally, if $u$ and $v$ belong to components of size $1$, then we must place them in a cluster different from $U$ and $V$.
	Note that in such case, a $C_1$-type cluster $W$ exists, as otherwise no component respecting partitioning exists. In this case \TAlg performs one swap, i.e., it exchanges the nodes $u$ and $v$ with any two nodes of $W$.
\end{proof}

@@

	%Let $A^I$ be the cost of (extra-cluster) communication incurred in this phase by \TAlg between pairs that belonged to the single component at the end of the phase.
	%Let $A^E$ be the cost of (extra-cluster) communication incurred in this phase between the nodes that belong to different components at the end of this phase.
	Let $\TAlg(M)$ be the cost of migrations performed by \TAlg in this phase.
	\TAlg performs at most $2 \ell$ component merge operations ---
	exceeding this number would mean that a~component of size $4$ exists, and the phase should have ended already.
	We bound the cost of a each repartition by Lemma~\ref{lem:1req}, to obtain $\TAlg(M) \leq 8\alpha\cdot\ell$.
	%Together with Lemma~\ref{lem:1req}, this allows us to bound the cost of migrations, $\TAlg(M) \leq 6\cdot\alpha\cdot\ell$.

@@

	Now we estimate the cost of $\OPT(\sigma_E)$ while accounting savings from migrations.
	By performing a swap of nodes $(u,v)$, $\OPT$ collocates $u$ with two nodes $u', u''$, and $v$ with two nodes $v'$, $v''$.
	This may allow serving requests between $(u,u')$, $(u,u'')$, $(v,v')$ and $(v,v'')$ for free afterwards.
	As $\sigma_E$ consists of requests to unsaturated pairs, and it accounts only for external requests, there are at most $\alpha-1$ requests between each of these pairs.
	By performing a single swap that costs $2\alpha$, $\OPT$ may avoid paying the remote serving costs for at most $4 (\alpha - 1)$ requests from $\sigma_E$
	Hence, for serving $\sigma_E$, $\OPT$ pays at least $\OPT(\sigma_E) + \OPT(M) \geq W \cdot \frac{2\alpha}{4 (\alpha-1)}\geq |\sigma_E| / 2 - 2 \alpha \cdot \ell$.

@@

While are bounds are tight, there are several interesting avenues for future research.
In particular, we have so far focused on deterministic algorithms, and it would be interesting to study the power of randomization in this context.
On the practical side, it would also be interesting to study our algorithms empirically, under realistic workloads.h

>>>>

\begin{abstract}
	Distributed   applications,  including  batch  processing, streaming, scale-out databases,
	or machine learning, generate a significant amount of network traffic.
	By collocating frequently communicating nodes (e.g., virtual machines) on the same clusters (e.g., server or rack), we can reduce the network load and  improve application performance. 
	However, the communication pattern of different applications is often unknown a priori and may change over time, hence it needs to be learned in an online manner.
	%
	This paper revisits the online 
	balanced repartitioning problem 
	(introduced by Avin et al.~at DISC 2016)
	that asks for an algorithm that strikes
	an optimal tradeoff between the benefits
	of collocation (i.e., lower network load) 
	and its costs (i.e., migrations). 
	%
	Our first contribution is a significantly improved
	lower bound of $\Omega(k\cdot \ell)$ on the
	competitive ratio, where $\ell$ is the number
	of clusters and $k$ is the cluster size,
	even for a scenario in which the communication
	pattern is static and can be perfectly partitioned;
	we also provide a \mahmoud{asymptotically?} tight upper bound 
	of $O(k\cdot \ell)$ for this scenario.
	We contribute a tight upper bound
	of $\Theta(\ell)$ for $k=3$,
	for the general model in which the
	communication pattern can change arbitrarily
	over time.
	We improve the result for $k=2$ by providing a strictly $6$-competitive upper bound for the general model.
	In contrast to most prior work, our algorithms respect all capacity constraints and do not require resource augmentation.

@@

The popularity of data-centric, distributed applications has led to an explosive growth of network traffic, especially in data centers \cite{roy2015inside,singh2015jupiter}.
The performance of these distributed applications often critically depends on the underlying network \cite{mogul2012we}, and efficient operation of these networks is important.
At the same time, distributed systems are often highly virtualized today, and provide interesting new opportunities for resource optimization.
In particular, it has become possible to operate data centers in a more demand-aware manner: 
by dynamically migrating nodes (e.g., virtual machines) which communicate frequently topologically closer to each other, network traffic can be reduced significantly.  
However, migrations should be used moderately, as migrations also entail overheads.

@@

An algorithm serves a communication request between two nodes
either \emph{locally} at cost~0
if they are collocated,
or \emph{remotely} at cost~1
if they are located in different clusters.
We refer to these two types of requests as \emph{internal}
and \emph{external} requests, respectively.
Before serving a request,
an online algorithm may perform a \emph{repartition},
%(i.e., \emph{reconfigure}).
i.e.,
it may move (``migrate'') some nodes into clusters different from their current clusters, while respecting the capacity of every cluster. 
Afterward, 
the algorithm serves the  request.
The cost of migrating a node from one cluster to another
is~$\alpha \in \mathbb{Z}^+$.
For any algorithm $\ALG$,
its cost,
denoted by $\ALG(\sigma)$,
is the total cost of communications and
the cost of migrations performed by $\ALG$ while serving the sequence $\sigma$.

@@

The two works closest to ours are by Avin et al. (on the general partitioning model) \cite{repartition-disc, sidma-arxiv} and by Henzinger et al. (on the learning model) \cite{sigmetrics19_partitioning}.
However, the focus of these papers is primarily on models with resource augmentation: the online algorithm can use slightly larger clusters than the offline algorithm.  
Avin et al. actually showed that their lower bound $\Omega(k)$ holds even for a significant resource augmentation, and they provided an algorithm with the competitive ratio $O(k \log k)$ using the $(2+\epsilon)$-augmented cluster capacity.
Their ratio is independent of $\ell$, which is impossible without significant resource augmentation.

@@

In this section, we consider the learning variant of Dynamic Balanced Graph Partitioning.
For this setting, we show a surprisingly high lower bound of $\Omega(k \cdot \ell)$.
The lower bound can be easily translated to the partitioning model (studied in Section~\ref{sec:part}) with a~finite request sequence.
At the end of this section, we discuss an asymptotically optimal upper bound.

@@

A~\emph{ground set} is a group of nodes that repeatedly communicate when an algorithm splits them (places at different clusters).
Their role is to ensure that the algorithm maintains a~partition of ground sets into clusters.
We emphasize that ground sets are revealed only upon their nodes are split.
We say that a ground set is \emph{trivial} if it contains exactly one node.
We say that the node that belongs to trivial a ground set is \emph{isolated}.

The adversary constructs ground sets depending on the choices of a deterministic online algorithm.
We emphasize that the algorithm does not know ground sets before they are revealed.
Once we construct a ground set, it lasts until the end of the input sequence.

@@

\begin{proof}
	Fix any online algorithm \ALG{}.
	Let $I(C)$ denote the cluster where nodes of the ground set $C$ are located in the initial configuration.
	If $C$ contains nodes that originate from different clusters, $I(C)$ is undefined.
	Throughout the construction, we maintain ground sets for each node.
	Initially, all nodes are isolated (i.e., in ground sets of size $1$).
	First, we create a ground set $B$ of $k-1$ nodes on an arbitrarily chosen cluster (the nodes  are already collocated).
	%We never dismiss any constructed ground set.
	Each cluster hosts exactly $k$ nodes, and in any configuration, a single isolated node resides in $S$.
	We say that this isolated node that accompanies the ground set $B$ is \emph{special}.
	We denote the first special node by $x_0$.

@@

	We claim that expanding ground sets in this way can continue as long as at least $\ell$ isolated nodes exist.
	Fix the configuration just before the creation of a ground set with $x_i$ for some $i$.
	We show that if at least $\ell$ isolated nodes exist (including $x_i$), then there exists a~perfect partition of ground sets after expanding a ground set with $x_i$.
	To show this claim, we consider a configuration, where the largest ground set of each cluster originates from it (excluding the ground set $\{x_0, y_0\}$).
	In this configuration, each cluster contains:  at most one ground set larger than $1$, possibly some isolated nodes, and possibly the ground set $\{x_0, y_0\}$.
	If $I(x_i)$ contains an isolated node, then we simply obtain a~perfect partition by swapping it with $x_i$.
	In the following, we assume that no isolated nodes exist on $I(x_i)$.
	The largest ground set of $I(x_i)$ is at most $k-1$, because $x_i$ is on $S$.
	Therefore $\{x_0, y_0\}$ is present on $I(x_i)$, and we must evict it to make space for $x_i$.
	To find the cluster for $\{x_0, y_0\}$, observe that we have at most $\ell-1$ clusters with isolated nodes, and we have at least $\ell$ isolated nodes --- hence a~cluster with two isolated nodes exists.
	To find a feasible partition, we swap $\{x_0, y_0\}$ with these two collocated isolated nodes, and then we swap an isolated node from $I(x_i)$ with $x_i$.

@@

\noindent
\textbf{General partitioning problem.}
Every lower bound for the learning problem constitutes a converging lower bound for the partitioning problem as the input sequence grows to infinity.
However, our construction allows for an exact lower bound for partitioning with a~finite input sequence.
To this end, we issue only enough requests until the algorithm collocates the nodes of each ground set (if it never does, we terminate after it pays $k\cdot \ell$).
The construction is oblivious to the choice of the reconfiguration cost $\alpha$: we compare the 	number of node exchanges of \ALG and \OPT.

@@

Finally, we note the possibility of improvement. The algorithm CREP~\cite{repartition-disc} requires $(2+\epsilon)$-augmentation to guarantee the competitive ratio independent of $\ell$.

@@

We say that a component is \emph{trivial} if it contains exactly one node,
and the node in this component is \emph{isolated}.
We say that a~partition is \emph{component-respecting}
if the nodes belonging to the same component are always collocated.

The algorithm in this section and the algorithm for $k=3$ (cf. Section~\ref{sec:k3}) are modified versions of the algorithm DET from \cite{repartition-disc}.
The difference is in the choice of partition after a component merge.
In DET, the partition was arbitrary.
In the algorithm for the learning model, we choose the partition closest to the initial partition.
\maciek{component respecting?}
In the algorithm from Section~\ref{sec:k3}, we choose the partition closest to the current partition (a repartition of minimum cost).

@@

With each repartitioning,
\PPL moves to a \maciek{component-respecting} partition that minimizes the distance to the initial partition $P_I$.
As a result,
\PPL never moves to a~partition that is more than $\Delta^*$  node migrations away from $P_I$.
This invariant latter ensures us that \PPL does not pay too much while recovering $P_F$.
We emphasize that a repartitioning by \PPL replaces the current partition $P$ with a~perfect partition closest to $P_I$.
Meaning, it never moves to a partition beyond the distance $\Delta^*$.

@@

\begin{lemma}	\label{lemma:rebalancecost}
	The cost of each repartitioning by \PPL is $2\cdot\OPT$.
\end{lemma}
\begin{proof}
	Let $P_i$ denote the partition of \PPL immediately after serving $\sigma_{i}$.
	Consider the repartitioning that transforms 
	$P_{t-1}$ to $P_t$ upon the request $\sigma_t$.
	Let $M \subset V$ denote the set of nodes that migrate during this process.
	Let $M^-$ and $M^+$ denote the subset of nodes that, respectively,
	enter or leave their initial cluster during the repartitioning.    
	Then,
	$M = M^+ \cup M^-$.
	Since at least $|M^-|$ nodes are not in their initial cluster before the repartitioning (i.e., in $P_{t-1}$),
	the distance before the repartitioning is $\Delta(P_{t-1}) \geq | M^-|$.
	Analogously,
	the distance afterward is $\Delta(P_{t}) \geq | M^+|$.
	Thus,
	$|M| \leq \Delta(P_{t-1}) + \Delta(P_{t})$.
	By Property \ref{prop:dist<OPT},
	$\Delta(P_{t-1})  \leq \Delta^*$ and
	$\Delta(P_{t}) \leq \Delta^*$.
	Since $\Delta^* \leq \OPT$,
	we obtain	 $|M| \leq 2\cdot\OPT$.
\end{proof}

@@

Now  we discuss the general online
model where the request sequence
can be arbitrary.
In Section~\ref{sec:k3}, we show an $O(k \cdot \ell)$-competitive algorithm for $k=3$ using the classic \emph{rent-or-buy} approach~\cite{karlin-ski-rental}.
Prior to this section, we showed a lower bound of $\Omega(k \cdot \ell)$  that holds for the general model, hence the result from this section is asymptotically optimal (cf. Section~\ref{sec:lowerbound}).
Furthermore, in Section~\ref{sec:k2}, we show a strictly $6$-competitive algorithm for $k=2$.

@@

	If either $U$ or $V$ is of type $C_1$, then this cluster can fit the merged component, and the reconfiguration is local within $U$ and $V$.
	In this case, a local repartition is possible (within $U$ and $V$), for a cost of at most $2$ swaps.
	Note that it is impossible that either $U$ or $V$ is of type $C_3$, as otherwise, a component of size $3$ participates in a merge.
	After the merge, we would have a component of size at least $4$, and in this case, no component respecting partitioning exists, a contradiction.

	It remains to consider the case where both $U$ and $V$ are of type $C_2$. Note that $(u,v)$ cannot both belong to a component of size $2$, as this would mean that \TAlg has the component of size $4$, a contradiction with the case assumption that the component respecting partitioning exists. 
	Otherwise, if either $u$ or $v$ belongs to a component of size $2$, then it suffices to exchange components of size $1$ between $U$ and $V$.
	Finally, if $u$ and $v$ belong to components of size $1$, then we must place them in a cluster different from $U$ and $V$.
	Note that in such case, a $C_1$-type cluster $W$ exists, as otherwise no component respecting partitioning exists. In this case, \TAlg performs one swap, i.e., it exchanges the nodes $u$ and $v$ with any two nodes~of~$W$.
\end{proof}

@@

	%Let $A^I$ be the cost of (extra-cluster) communication incurred in this phase by \TAlg between pairs that belonged to the single component at the end of the phase.
	%Let $A^E$ be the cost of (extra-cluster) communication incurred in this phase between the nodes that belong to different components at the end of this phase.
	Let $\TAlg(M)$ be the cost of migrations performed by \TAlg in this phase.
	\TAlg performs at most $2 \ell$ component merge operations ---
	exceeding this number would mean that a~component of size $4$ exists, and the phase should have ended already.
	We bound the cost of each repartition by Lemma~\ref{lem:1req}, to obtain $\TAlg(M) \leq 8\alpha\cdot\ell$.
	%Together with Lemma~\ref{lem:1req}, this allows us to bound the cost of migrations, $\TAlg(M) \leq 6\cdot\alpha\cdot\ell$.

@@

	Now we estimate the cost of $\OPT(\sigma_E)$ while accounting savings from migrations.
	By performing a swap of nodes $(u,v)$, $\OPT$ collocates $u$ with two nodes $u', u''$, and $v$ with two nodes $v'$, $v''$.
	This may allow serving requests between $(u,u')$, $(u,u'')$, $(v,v')$ and $(v,v'')$ for free afterward.
	As $\sigma_E$ consists of requests to unsaturated pairs, and it accounts only for external requests, there are at most $\alpha-1$ requests between each of these pairs.
	By performing a single swap that costs $2\alpha$, $\OPT$ may avoid paying the remote serving costs for at most $4 (\alpha - 1)$ requests from $\sigma_E$
	Hence, for serving $\sigma_E$, $\OPT$ pays at least $\OPT(\sigma_E) + \OPT(M) \geq W \cdot \frac{2\alpha}{4 (\alpha-1)}\geq |\sigma_E| / 2 - 2 \alpha \cdot \ell$.

@@

While our bounds are tight, there are several interesting avenues for future research.
In particular, we have so far focused on deterministic algorithms, and it would be interesting to study the power of randomization in this context.
On the practical side, it would also be interesting to study our algorithms empirically, under realistic workloads.
